---
title: "Cheatsheet"
author: "Massimiliano"
date: "3/6/2017"
output: 
  html_document:
    number_sections: yes
    toc: yes
    css: style.css
---

# To Start

These notes contain the basic codes to start writing economic equations and formulas in Tex language.

Each code is included in an easy and more complex example.

# Greek letters

To write Greek letters the symbol of "\\" (backslash) is followed by the name of the Greek letter. 

Note that the symbol of "\\" is one of the main commands to write codes. 

For example, some of the most used Greek letters in Economics:

```{r, include = FALSE}
source("utils.R")
```


```{r, echo = FALSE}
greek_letters = c('alpha', 'beta', 'gamma', 'theta', 'sigma', 
                  'rho', 'lambda', 'phi', 'psi', 'epsilon')
knitr::kable(math_table(greek_letters))
```

For the corresponding capital letter, we just write in the same way but with initial capital letter.

Example

```{r, echo = FALSE}
capital_letters = c("Gamma", "Lambda", "Sigma", "Psi", 
                    "Delta", "Xi", "Upsilon", "Omega", 
                    "Theta", "Pi", "Phi")
knitr::kable(math_table(capital_letters))
```


#Parenthesis
We can write parenthesis and brackets just giving the input on the keyboard $()$ or $[]$. This does not work for braces.

Another way is to write parenthesis with the following notation:

```{r, echo = FALSE, results = "asis"}
s_paren = paste(
  "\\left(",
  "\\right) \\\\",
  "\\left[",
  "\\right]", 
  sep = "\n") %>% two_column_layout
```


that is writing the symbol of `\` (blackslash) followed by `\left` _before opening the parenthesis_ and _blackslash_ followed by `\right` _before closing the parenthesis_. 

In case of braces, we should use the following notation:

```{r, echo = FALSE, results = "asis"}
s_curly = paste(
  "\\left\\{", 
  "\\right\\}",
  sep = "\n") %>% two_column_layout
```

The difference with respect to parenthesis and brackets is a _backslash after left and right_.

Example:

```{r, echo = FALSE, results = "asis"}
s_complex_braces = paste(
"10 \\cdot ",
"  \\left\\{",
"    5 + \\left[",
"      3 + 2 \\cdot ",
"         \\left(",
"            9-1\\right)",
"        \\right]  ",
"  \\right\\}",
sep = "\n") %>% two_column_layout
```

The opportunity to write the full notation for parenthesis and brackets will be clear when dealing whit fractions. 

#Fractions
To write fractions the inputs to give are \textbf{\textit{backslash}}, \textbf{\textit{frac}} and then \textit{numerator and denominator written between two braces}. With an example will be clear.

Example:

```{r, echo = FALSE, results = "asis"}
s_frac = paste(
"\\frac{5}{2}",
sep = "\n")%>% two_column_layout

```


Example:

```{r, echo = FALSE, results = "asis"}
s_frac01 = paste(
"\\frac{\\alpha}{1-\\alpha} = \\phi",
sep = "\n")%>% two_column_layout

```



#Paranthesis & Fraction
If we write a fraction in parenthesis with the shortcut

```{r, echo=FALSE, results="asis"}
s_pfrac = paste(
 "( \\frac{5}{2})", 
sep = "\n")%>% two_column_layout

```

the parenthesis do not fit the fraction.

If we write with full notation

```{r, echo=FALSE, results="asis"}
s_pfrac01 = paste(
  "\\left(",
  "\\frac{5}{2}",
  "\\right)",
sep = "\n")%>% two_column_layout

```

we will have an appropriate result.

Example:

```{r, echo=FALSE, results="asis"}
s_pfrac02 = paste(
  "\\frac{Y}{L} =",
  "\\left(",
  "\\frac{K}{L}, 1",
  "\\right)",
sep = "\n")%>% two_column_layout

```



#Exponential

To write the exponential we use the symbol `^` 

```{r, echo=FALSE, results="asis"}
s_expo = paste(
"2^5",
sep = "\n")%>% two_column_layout

```


In case of operation in the exponential, we need to _use braces after_ `^` 


Example:

```{r, echo=FALSE, results="asis"}
s_expo01 = paste(
 "F(K, L) = ",
 "K^\\alpha L^\\beta",
sep = "\n")%>% two_column_layout

```


Example:

```{r, echo=FALSE, results="asis"}
s_expo02 = paste(
  "F(K, L) =",
  "\\left[",
  "\\gamma K^\\frac{\\sigma - 1}{\\sigma}",
  " + (1 - \\gamma)L^\\frac{\\sigma - 1}{\\sigma}",
  "\\right]^\\frac{\\sigma}{\\sigma - 1}",
sep="\n")%>% two_column_layout

```



#Subscript

To write subscript use the symbol `_`

```{r, echo=FALSE, results="asis"}
s_sub = paste(
  "t_0",
sep= "\n")%>% two_column_layout

```



We use _braces in case of operation in subscript_

Example

```{r, echo=FALSE, results="asis"}
s_sub01 = paste(
  "y_ty_{t-j}",
sep = "\n")%>% two_column_layout
```


#Derivative

Example:

```{r, echo=FALSE, results="asis"}
s_der = paste(
  "\\frac{dk}{k}",
sep="\n")%>% two_column_layout
```

To write partial differentiation, write the symbol 
`\` followed by `\partial` in the fraction notation

Example:

```{r, echo=FALSE, results="asis"}
s_der01 = paste(
  "\\frac{",
  "\\partial F}{",
  "\\partial K}",
  "r + \\delta",
sep="\n")%>% two_column_layout
```


#Lagrangian

The symbol of Lagrangian is written as follows:

```{r, echo = FALSE, results="asis"}
s_lag = paste(
  "\\mathcal L",
sep = "\n")%>% two_column_layout

```



#Summation

The code to write summation is the following: symbol of `\` followed by _sum_, `_` and _braces with lower bound of summation_ followed by the symbol `^` _before the brace with the upper bound of summation_. 

Example:

```{r, echo=FALSE, results="asis"}
s_sum = paste(
  "\\sum_{t_0}^{\\infty}",
sep = "\n")%>% two_column_layout
```

An example summing up what we saw until now:

```{r, echo=FALSE, results="asis"}
s_sum01 = paste(
  "\\mathcal L =",
  "\\sum_{t=0}^{\\infty}",
  "\\left(",
  "\\frac {1}{1 + \\rho}",
  "\\right)^{t} u(c_t)",
  "\\frac{L_t}{H}",
  " + \\lambda",
  "\\left[",
  "S_0 + \\sum_{t=0}^{\\infty}",
  "R_t^{-1}w_t",
  "\\left(",
  "\\frac{L_t}{H}",
  "\\right)",
  "- \\sum_{t=0}^{\\infty} R_t^{-1}c_t",
  "\\left(",
  "\\frac{L_t}{H}",
  "\\right)",
  "\\right]",
sep="\n")%>% two_column_layout
```



#Limit

The code for limit is written as follows:

```{r, echo = FALSE, results="asis"}
s_lim = paste(
  "\\lim_{t\\to\\infty}",
sep = "\n")%>% two_column_layout
```


Example: 

```{r, echo=FALSE, results="asis"}
s_lim01 = paste(
  "\\lim_",
  "{t\\to\\infty}",
  "R_t^{-1}S_t = 0",
sep = "\n")%>% two_column_layout
```

Example

```{r, echo = FALSE, results="asis"}
s_lim02 = paste(
  "e^{",
  "lim_{\\gamma\\to0}",
  "\\frac{1}{\\gamma}ln",
  "\\left[",
  "\\alpha K^\\gamma + (1 - \\alpha)L^\\gamma",
  "\\right]}",
sep = "\n")%>% two_column_layout
```


#Integrals

```{r, echo = FALSE, results="asis"}
s_int = paste(
  "\\int_0^\\infty",
sep = "\n")%>% two_column_layout
```


Example

<div class="twocol">
<div class="left">
```
$$
\int_0^\infty
  \left[
    r(t)e^{-R(t)}S(t) 
    + e^{-R(t)} w(t) 
    \frac{L(t)}{H}
    -e^{-R(t)}c(t) 
    \frac{L(t)}{H} 
  \right]
  dt
$$
```
</div><!-- .left -->
<div class="right">
$$
\int_0^\infty\left[r(t)e^{-R(t)}S(t) + e^{-R(t)} w(t) \frac{L(t)}{H}-e^{-R(t)}c(t) \frac{L(t)}{H} \right]dt
$$
</div><!-- .left -->
</div>


#Matrix

The followin example shows how to write a matrix.

Note that `\bmatrix` stands for _bracket matrix_ while `\pmatrix` stands for _parenthesized matrix_. Other kinds are available.

Moreover, `\vdots` stands for _vertical dots_ while `\ddots` for _diagonal dots_. Other version of dots includes `\cdots` that stands for _central dots_, used also as multiplication sign.

<div class="twocol">
<div class="left">
```
$$
A =
\begin{bmatrix}
  a_{11} & a_{12} & a_{13} & \dots & a_{1n} \\
  a_{21} & a_{22} & a_{23} & \dots & a_{2n} \\
  \vdots & \vdots & \vdots & \vdots & \vdots\\
  a_{m1} & a_{m2} & a_{m3} & \dots & a_{mn} \\
\end{bmatrix}
\ne
B =
\begin{pmatrix}
 b_{11} & b_{12} & b_{13} & \cdots & b_{1n} \\
 b_{21} & \ddots & b_{23} & \cdots & b_{2n} \\
 ...&...&... & \ddots &... \\
 b_{m1} & b_{m2} & b_{m3} & \dots & b_{mn} \\
\end{pmatrix}
$$
```
</div><!-- .left -->
<div class="right">
$$
A =
\begin{bmatrix}
  a_{11} & a_{12} & a_{13} & \dots & a_{1n} \\
  a_{21} & a_{22} & a_{23} & \dots & a_{2n} \\
  \vdots & \vdots & \vdots & \vdots & \vdots\\
  a_{m1} & a_{m2} & a_{m3} & \dots & a_{mn} \\
\end{bmatrix}
\ne
B =
\begin{pmatrix}
 b_{11} & b_{12} & b_{13} & \cdots & b_{1n} \\
 b_{21} & \ddots & b_{23} & \cdots & b_{2n} \\
 ...&...&... & \ddots &... \\
 b_{m1} & b_{m2} & b_{m3} & \dots & b_{mn} \\
\end{pmatrix}
$$
</div><!-- .left -->
</div>

#Equalities and Inequalities

```{r, echo = FALSE}
equalities = tribble(
  ~desription,                ~cmd,    ~asis,
  "Equality or assignment",   "=",     TRUE,
  "Inequality",               "neq",   FALSE,
  "Less than",                "<",     TRUE,
  "Less than or equal to",    "leq",   FALSE,
  "Greater than",             ">",     TRUE,
  "Greater than or equal to", "geq",   FALSE,
  "Approximately equal to",   "simeq", FALSE)

equalities = equalities %>% 
  by_row(function(x) put_codechunk(x$cmd, x$asis), 
         .collate = "rows", .to = "latex") %>% 
  by_row(function(x) put_result(x$cmd, x$asis), 
         .collate = "rows", .to = "result") %>% 
  select(-asis, -cmd)

knitr::kable(equalities)
```

#Subset and Functions

```{r, echo = FALSE}
set_operations = tribble(
  ~desription,            ~cmd,
  "A is included in B",   "A \\subset B",
  "A includes B",         "A \\supset B",
  "A intersection B",     "A \\cap B",
  "A union B",            "A \\cup B", 
  "x is in A",            "x \\in A",
  "A contains x",         "A \\ni x",
  "x is not in A",        "x \\notin A",
  "for some x",           "\\exists x",
  "for any x",            "\\forall x",
  "tends to / maps into", "\\to",
  "gets",                 "\\gets",  # I've never used this...
  "A implies B",          "A \\implies B")

set_operations = set_operations %>% 
  by_row(function(x) put_codechunk(x$cmd, TRUE), 
         .collate = "rows", .to = "latex") %>% 
  by_row(function(x) put_result(x$cmd, TRUE), 
         .collate = "rows", .to = "result") %>% 
  select(-cmd)

knitr::kable(set_operations)
```


